---
title: "ice_meteor"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(here)
library(ggplot2)
library(readr)
```

-   In a single R Markdown (or Quarto) document:

    -   Read in both files (ice cover duration and meteorological data). Note that the ice cover data has some NA values stored as -999.

    ```{r}

     meteor<- read_csv(here("data", "ntl20_v6.csv"), 
                          na = c("-999", "NA"))

     ice_cover<- read_csv(here("data", "ntl33_v7.csv"), 
                       na = c("-999", "NA"))

    ```

    -   Create a visualization that compares ice duration across different lakes in the **Madison Lake Area**. Include both the actual ice duration values *and* some representation of summary statistics for each lake. Careful if you use a boxplot over a jitter or beeswarm - what values might be doubly represented? ***Can you find an argument that would help you remove those from the boxplot?***

    ```{r}

    ice_across_lakes_plot <- ggplot(data = ice_cover, aes(x= lakeid, y = ice_duration)) + geom_jitter() + geom_boxplot()
    ice_across_lakes_plot
    ```

    -   Find the mean ice cover duration in Madison Lake Area by year (i.e., you\'ll only be grouping by the year variable here). Understanding check: what are you finding the mean of here? Why are there multiple observations for each year?)

    ```{r}
    mean_ice_by_yr <- ice_cover %>% 
      group_by(year4) %>% 
      summarize(mean_ice_duration_by_yr= mean(ice_duration, na.rm = TRUE))
    mean_ice_by_yr
    ```

    -   Create an exploratory visualization of mean ice cover duration by year for the Madison Lake Area. Add a sentence or two below this exploratory graph describing the overall trend(s) you observe.

    ```{r}
    mean_ice_duration_by_year <- mean_ice_by_yr %>% 
      ggplot(aes(x=year4, y=mean_ice_duration_by_yr))+
      geom_point()
    mean_ice_duration_by_year 
    ```

    -   Find the mean air temperature (using the daily average air temperature - adjusted) in Madison Lake Area **by year** using **only observations from winter months (December, January, February)**

    ```{r}
    mean_air_temp_winter_table <- meteor %>% filter(month %in% c(12, 1, 2)) %>% 
      group_by(year4) %>% 
      summarize(mean_temp= mean(ave_air_temp_adjusted, na.rm = TRUE))
    mean_air_temp_winter_table
    ```

    -   Create an exploratory visualization of mean winter temperatures for the Madison Lake Area. Add a sentence or two below this exploratory graph describing the overall trend(s) you observe.

    ```{r}
    mean_wint_temps_Madison_plot <- mean_air_temp_winter_table %>% 
      ggplot(aes(x= year4, y=mean_temp))+geom_point()
    mean_wint_temps_Madison_plot
    ```

-   

    -   Join the mean winter air temperatures to the mean ice cover duration data you found above

        ```{r}
        wint_airtemp_ice_dur <- left_join(mean_air_temp_winter_table , mean_ice_by_yr, by = year4)
        ```

-   

    -   Create an exploratory scatterplot of mean winter temperatures versus mean ice duration. Add a sentence or two below this exploratory graph describing the overall trend(s) you observe.

    -   Explore thaw dates since 1970. Create a subset of the ice cover duration data since 1970 (ok to include 1970), then convert the `ice_off` column to a `Date`. Use `lubridate::yday()` to pull the numeric *day of the year* from that Date. Create an exploratory plot of the ice off day (numeric day-of-the-year) from 1970 - present. Add a sentence or two below this exploratory graph describing the overall trend(s) you observe.
